Robust Bayesian Analysis of Causal Inference Problems

### 2022-10-29 JASA-template.tex
more cleaning

\subsection{Regression Model}
Regression methods are widely used in causal effect estimation. The
main idea behind these regression methods is to remove the
correlation between the treatment indicator and the error term
\cite{winship99,HECKMAN1985}. To do so, we rely on $p$ different observed quantities
or predictors denoted by $X\coloneqq$ $[X_1$, \dots, $X_n]^T$. Now, let
$\beta \coloneqq (\beta_1$, \dots, $\beta_p)$ denotes the vector of regression
coefficients. Then we can define a linear model for the outcome
so that
\begin{equation}
    Y_i = \beta_{T} T_i + \beta_0 + X\beta + \epsilon_i
\end{equation}
where $\epsilon_i\sim \mathcal{N}(0, \sigma^2)$. Clearly, when
the underlying true outcome model is linear,
\begin{equation}
    \delta = \beta_{T}\quad\text{and hence } 
    \hat{\delta} - \delta = \hat{\beta}_T - \beta_T.
\end{equation}
\begin{figure}
	\centering
	\begin{tikzpicture}[params/.style={circle, draw=green!60, fill=green!5, very thick, minimum size=7mm}]
	\node (1) at (0,0) {Treatment ($T$)};
	\node (2) [right = of 1] {Outcome ($Y$)};
	\node (4) [below = of 1] {Input ($X$)};

@@

\citet{koch2020} suggested the use of a probit model to specify the conditional probability that
subject $i$ receives the treatment. That is,
\begin{align}
    P(T_i=1\mid X_i) = \Phi(X_i^T\gamma).
\end{align}
Therefore, we can define intermediate variables $T_i^*$ so that, $T_i=1$ if $T_i^*>0$ and
$T_i=0$ if $T_i^*\le0$. Now, following the approach of \citet{koch2018}, we define an adjusted
output vector $W\coloneqq(Y, T^*)^T$ and corresponding $2n\times(2p+1)$ dimensional design matrix
\begin{align}
    Z &=
    \begin{bmatrix}
     X_O & 0 \\
     0 & X
    \end{bmatrix},
\end{align}
where, $X_O = [T, X]$. Then, assuming Gaussian error term, we have the following likelihood function
\begin{align}
W\mid Z, \alpha, \beta, \gamma, \sigma^2 \sim\normal\left(Z\nu, \sigma^2\mathbf{I}_{2n}\right),
\end{align}
where $\nu = (\alpha, \beta, \gamma)^T$.

>>>>

\subsection{Regression Model}
Regression methods are widely used in causal effect estimation. The
main idea behind these regression methods is to remove the
correlation between the treatment indicator and the error term
\cite{winship99,HECKMAN1985}. To do so, we rely on $p$ different observed quantities
or predictors denoted by $X\coloneqq$ $[X_1$, \dots, $X_n]^T$. Now, let
$\beta \coloneqq (\beta_1$, \dots, $\beta_p)$ denotes the vector of regression
coefficients. Then we can define a linear model for the outcome
so that
\begin{equation}
    Y_i = \beta_{T} T_i + \beta_0 + X_i\beta + \epsilon_i
\end{equation}
where $\epsilon_i\sim \mathcal{N}(0, \sigma^2)$. Clearly, when
the underlying true outcome model is linear,
\begin{equation}
    \delta = \beta_{T}\quad\text{and hence } 
    \hat{\delta} - \delta = \hat{\beta}_T - \beta_T.
\end{equation}
\begin{figure}
	\centering
	\begin{tikzpicture}[params/.style={circle, draw=green!60, fill=green!5, very thick, minimum size=7mm}]
	\node (1) at (0,0) {Treatment ($T$)};
	\node (2) [right = of 1] {Outcome ($Y$)};
	\node (4) [below = of 1] {Predictors ($X$)};

@@

In the presence of confounders we also need to consider the
association between the treatment indicators and the predictors.
\citet{koch2020} suggested the use of a probit link function to 
construct the regression model. This way, we can
specify the conditional probability that subject $i$ receives the treatment through a linear model. 
That is, for another vector of regression coefficients 
$\gamma\coloneqq(\gamma_1, \cdots, \gamma_p)$ we define
\begin{align}
    P(T_i=1\mid X_i) \coloneqq \Phi(\gamma_0+X_i\gamma)
\end{align}
where $\Phi(\cdot)$ denotes the cumulative distribution function
of a standard normal distribution.

This allows us to define intermediate latent variables 
\begin{equation}
	T_i^* = \gamma_0 + X_i\gamma +u_i
\end{equation}
where, $u_i\sim\mathcal{N}(0,1)$. Therefore, $T_i=1$ if $T_i^*>0$ and
$T_i=0$ if $T_i^*\le0$.

Now, following the approach of \citet{koch2018}, we define an adjusted
output vector $W\coloneqq(Y, T^*)^T$ and corresponding $2n\times(2p+1)$ dimensional design matrix
\begin{align}
    Z &=
    \begin{bmatrix}
     X_O & 0 \\
     0 & X
    \end{bmatrix},
\end{align}
where, $X_O = [T, X]$. Then, assuming Gaussian error term, we have the following likelihood function
\begin{align}
W\mid Z, \alpha, \beta, \gamma, \sigma^2 \sim\normal\left(Z\nu, \sigma^2\mathbf{I}_{2n}\right),
\end{align}
where $\nu = (\alpha, \beta, \gamma)^T$.
